---
title: "Appendix"
author: "Seth Daly, Duc Vu, Bich Ha Nguyen, Ngoc Xuan Dung Nguyen"
date: "11/22/2021"
output:
  word_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

#pdf_document

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Import libraries

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "1"}
library(MASS)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidyr)
library(rqdatatable)
library(lubridate)
library(caret)
library(xts)
library(forecast)
library(tsbox)
library(TSstudio)
library(knitr)  
library(kableExtra) # formatting table
library(gtools)
library(TSA)
library(tseries)
library(gridExtra)
library(ggpubr)
```

# Data Cleaning

## Read In

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "2"}
droughtData <- read_csv("SPI.csv",
      col_names = c('SPI', 'Date', 'D0', 'D1', 'D2','D3', 'D4', 'SPI2', 'W0', 'W1', 'W2', 'W3','W4'), skip=1)
droughtData <- droughtData %>%
mutate(Date = ymd(substring(droughtData$'Date', 3)))
droughtData <- droughtData %>% filter(SPI > 0)
droughtData_D0 <- ts(droughtData$D0, start=c(1895, 9), end=c(2022, 3), freq=12)



### From 1895/09 to 1956/12
droughtData_1956 <- ts(droughtData_D0, start=c(1895, 9), end=c(1956, 12), freq=12)


### From 1957/01 to 2022/03
droughtData_2022 <- ts(droughtData_D0, start=c(1957, 1), end=c(2022, 3), freq=12)



### Train/Val Split from 1895/09 to 1956/12
train_1956 <- window(droughtData_1956, start = c(1895, 1), end=c(1955, 12))
val_1956 <- window(droughtData_1956, start = c(1956, 1), end=c(1956, 12))



### Train/Val Split from 1957/01 to 2022/03
train_2022 <- window(droughtData_2022, start = c(1957, 1), end=c(2020, 12))
val_2022 <- window(droughtData_2022, start = c(2021, 1), end=c(2022, 3))
```

# EDA

## Variable meaning (important)
Dry spell percentiles and corresponding
USDM categories: 
D0 = At Least Abnormally Dry, 
D1 = At Least Moderate Drought, 
D2 = At Least Severe Drought, 
D3 = At Least Extreme Drought, 
D4 = At Least Exceptional Drought. 

Note, categories are not mutually exclusive:

D0 + D1 + D2 + D3 + D4 != total amount of droughts!

D0 is total amount of droughts

All these values are percent of the country suffering this issue.

## Structure
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "3"}
head(droughtData_D0)
tail(droughtData_D0)
dim(droughtData_D0)
```


## Old Data

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "4"}
dim(droughtData_1956)
head(droughtData_1956)
tail(droughtData_1956)

autoplot(droughtData_1956)
```

## New Data

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "5"}
dim(droughtData_2022)
head(droughtData_2022)
tail(droughtData_2022)

autoplot(droughtData_2022)
```

## All Data
```{r}
## Autoplot of all data
ggplot2::ggplot(data=droughtData) +  
geom_line(aes(x=Date, y=D0), size=0.4) +
  	scale_color_manual(name="", values=c("Blue")) + 
  	labs(x="Year", y="Percentage")+
 	ggtitle("Percentage of Abnormal Drought in US (1895-2022)") + 
  	theme(legend.position="bottom")

## Dickey Fuller test
adf.test(droughtData$D0)

## Trend & Seasonality + Remainder
       autoplot(stl(droughtData_D0, s.window = "periodic"))

       ## The distribution of D0 values
       ggplot(data = droughtData, aes(x=D0)) + 
geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) + 
labs(x="Percentage of Abnormal Drought", y="Frequency") +
geom_vline(aes(xintercept=mean(D0)), color="blue", linetype="dashed", size=1)
```

# Holt Winters Modeling

## Holt Winters Modeling (Old D0)

### Model Creation

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "6"}
D0TS.full.old <- droughtData_1956
D0TS.full <- droughtData
D0TS.full.new <- droughtData_2022
D0TS.test.old <- val_1956
D0TS.train.old <- train_1956
D0TS.test.new <- val_2022
D0TS.train.new <- train_2022

D0.exp.old <- HoltWinters(D0TS.train.old, beta=FALSE, gamma=FALSE)
D0.trend.old <- HoltWinters(D0TS.train.old, gamma=FALSE)           
D0.season.old <- HoltWinters(D0TS.train.old, beta=FALSE)      
D0.hw.old <- HoltWinters(D0TS.train.old)  
```

### Model Predictions

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "7"}
D0.for.exp.old <- forecast(D0.exp.old, h=length(D0TS.test.old))
D0.for.trend.old <- forecast(D0.trend.old, h=length(D0TS.test.old))
D0.for.season.old <- forecast(D0.season.old, h=length(D0TS.test.old))
D0.for.hw.old <- forecast(D0.hw.old, h=length(D0TS.test.old))
```

### Visual Model Comparison

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "8"}
autoplot(window(D0TS.train.old, start=c(1950,1))) + 
  autolayer(D0.for.exp.old, PI=FALSE, series="Exponential") +
  autolayer(D0.for.trend.old, PI=FALSE, series="Holt Winters (Trend Only)") +
  autolayer(D0.for.season.old, PI=FALSE, series="Holt Winters (Seasonality Only) ") +
  autolayer(D0.for.hw.old, PI=FALSE, series="Holt Winters") +
  autolayer(D0TS.test.old, series="True Drought Percentage") +
  labs(x="Year", y="Percentage of Country Experiencing Drought") +
  theme_bw()
```

### Compare Model Accuracy

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "9"}
tab <- rbind(accuracy(D0.for.exp.old, D0TS.test.old)[2,2:3],
             accuracy(D0.for.trend.old, D0TS.test.old)[2,2:3],
             accuracy(D0.for.season.old, D0TS.test.old)[2,2:3],
             accuracy(D0.for.hw.old, D0TS.test.old)[2,2:3] )
rownames(tab) <- c("Exp Smoothing", "Holt (trend)", "HW (Season only)", "Holt Winters")
kable(tab)
```

From a combination of RMSE and MAE, we select the Holt Winters Model.

### Visual Final Model Comparison

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "10"}
autoplot(window(D0TS.train.old, start=c(1950,1))) + 
  autolayer(D0.for.hw.old, PI=TRUE, series="Holt Winter") +
  autolayer(D0TS.test.old, series="True Drought Percentage") +
  labs(x="Year", y="Percentage of Country Experiencing Drought") +
  theme_bw()
```

The Holt Winters model appears to be a fairly accurate measure of trend, but not of specific spikes and drops.  The prediction bands are extremely wide, which may indicate our lack of confidence in this specific prediction.

## Holt Winters Modeling (New D0)

### Model Creation

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "11"}
D0.exp.new <- HoltWinters(D0TS.train.new, beta=FALSE, gamma=FALSE)
D0.trend.new <- HoltWinters(D0TS.train.new, gamma=FALSE)           
D0.season.new <- HoltWinters(D0TS.train.new, beta=FALSE)      
D0.hw.new <- HoltWinters(D0TS.train.new)  
```

### Model Predictions

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "12"}
D0.for.exp.new <- forecast(D0.exp.new, h=length(D0TS.test.new))
D0.for.trend.new <- forecast(D0.trend.new, h=length(D0TS.test.new))
D0.for.season.new <- forecast(D0.season.new, h=length(D0TS.test.new))
D0.for.hw.new <- forecast(D0.hw.new, h=length(D0TS.test.new))

```

### Visual Model Comparison

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "13"}
autoplot(window(D0TS.train.new, start=c(2010,1))) + 
  autolayer(D0.for.exp.new, PI=FALSE, series="Exponential") +
  autolayer(D0.for.trend.new, PI=FALSE, series="Holt Winters (Trend Only)") +
  autolayer(D0.for.season.new, PI=FALSE, series="Holt Winters (Seasonality Only) ") +
  autolayer(D0.for.hw.new, PI=FALSE, series="Holt Winters") +
  autolayer(D0TS.test.new, series="True Drought Percentage") +
  labs(x="Year", y="Percentage of Country Experiencing Drought") +
  theme_bw()
```

### Compare Model Accuracy

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "14"}
tab <- rbind(accuracy(D0.for.exp.new, D0TS.test.new)[2,2:3],
             accuracy(D0.for.trend.new, D0TS.test.new)[2,2:3],
             accuracy(D0.for.season.new, D0TS.test.new)[2,2:3],
             accuracy(D0.for.hw.new, D0TS.test.new)[2,2:3] )
rownames(tab) <- c("Exp Smoothing", "Holt (trend)", "HW (Season only)", "Holt Winters")
kable(tab)
```

From a combination of RMSE and MAE, we select the Holt Winters (Season Only) model.

### Visual Final Model Comparison

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "15"}
autoplot(window(D0TS.train.new, start=c(2010,1))) + 
  autolayer(D0.for.season.new, PI=TRUE, series="Holt Winter (Season Only)") +
  autolayer(D0TS.test.new, series="True Drought Percentage") +
  labs(x="Year", y="Percentage of Country Experiencing Drought") +
  theme_bw()
```

# Cosine-Sine Model

## Part 1: Old Period 

### Examine Data

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "19"}
train_1956 %>%
  ggtsdisplay(xlab="Year",
              main="D0 Old data Drought")


train_1956 %>% diff() %>%
  ggtsdisplay(xlab="Year",
              main="Differenced D0 Old data Drought")
```

### A. Periodogram
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "20"}
f1956 = spec.pgram(train_1956, log="no", spans=c(9,9), detrend = TRUE)
f1956d = spec.pgram(diff(train_1956), log="no", spans=c(9,9), detrend = TRUE)
```

### identify frequency
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "21"}
del <- 12
spx1956 <- f1956d$freq/del
spy1956 <- 2*f1956d$spec
plot(spy1956~spx1956,xlab="frequency",ylab="spectral density",type="l")
#### frequencies: 0.04, 0.19, 0.25, 0.4, 0.5
abline(v=0.05, col="red") 
abline(v=0.155, col="blue")
abline(v=0.275, col="green")
abline(v=0.39, col="pink")
```

### Fit cosine-sine regression model
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "22"}
w1=0.05
w2=0.155
w3=0.275
w4=0.39
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "23"}
t= 1:length(diff(train_1956))
cs1=cos(2*pi*w1*t) 
si1=sin(2*pi*w1*t) 
cs2=cos(2*pi*w2*t) 
si2=sin(2*pi*w2*t)
cs3=cos(2*pi*w3*t) 
si3=sin(2*pi*w3*t) 
cs4=cos(2*pi*w4*t) 
si4=sin(2*pi*w4*t) 
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "24"}
fit1.1=lm(diff(train_1956)~t+cs1+si1+cs2+si2+cs3+si3+cs4+si4) 
summary(fit1.1)

fit1.2=lm(diff(train_1956) ~ si1-1) 
summary(fit1.2) 

resi1.1=fit1.2$residuals
temp1 <- ts(resi1.1, start=c(1895, 9), frequency=12)
temp1 %>% ggtsdisplay()
```

### Fit with ARIMA 

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "25"}
t= 1:length(train_1956)
si1=sin(2*pi*w1*t) 


fitp1.1 <- Arima(train_1956, order=c(2, 1, 2), xreg=cbind(si1))
fitp1.2 <- Arima(train_1956, order=c(2, 1, 0), xreg=cbind(si1))
fitp1.3 <- Arima(train_1956, order=c(0, 1, 2), xreg=cbind(si1))
fitp1.4 <- Arima(train_1956, order=c(1, 1, 1), xreg=cbind(si1))
fitp1.5 <- Arima(train_1956, order=c(2, 1, 1), xreg=cbind(si1))
fitp1.6 <- Arima(train_1956, order=c(1, 1, 2), xreg=cbind(si1))
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "26"}
getAIC <- function(fit) {
   c(fit$aic, fit$aicc, fit$bic)
}

tab <- rbind(getAIC(fitp1.1), getAIC(fitp1.2), getAIC(fitp1.3),
             getAIC(fitp1.4), getAIC(fitp1.5), getAIC(fitp1.6))
tab
colnames(tab) <- c("AIC", "AICc", "BIC")
rownames(tab) <- c("ARIMA(2, 1, 2)",
                   "ARIMA(2, 1, 0)",
                   "ARIMA(0, 1, 2)",
                   "ARIMA(1, 1, 1)",
                   "ARIMA(2, 1, 1)",
                   "ARIMA(1, 1, 2)")
tab %>% kbl()%>%
  kable_material(c("striped", "hover"))
```


### Model Accuracy
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "27"}
t1.test = 1:length(val_1956)
si1.t=sin(2*pi*w1*t1.test) 

d0_1.validation <- forecast(fitp1.1, xreg=cbind(si1.t))
acc1 <- accuracy(d0_1.validation, val_1956)
acc1
acc1[2, 2:3]

d0_1.validation2 <- forecast(fitp1.5, xreg=cbind(si1.t))
acc2 <- accuracy(d0_1.validation2, val_1956)
acc2
```

### Visual plot

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "28"}
autoplot(train_1956) +
  autolayer(fitted(fitp1.1))

autoplot(val_1956) + 
  autolayer(d0_1.validation)

autoplot(window(train_1956, start=c(1950,1))) + 
  autolayer(d0_1.validation, series="arima with cosine") + 
  autolayer(val_1956, series="True Percentage", size=1.5) +
  labs(x="Year", y="Percentage") +
  theme_bw()
```


### Forecast 3 more data points
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "29"}
t1_f= 1:length(droughtData_1956)
si1_f=sin(2*pi*w1*t1_f) 


fullfit1 <- Arima(droughtData_1956, order=c(2, 1, 2), xreg=cbind(si1_f))
summary(fullfit1)
checkresiduals(fullfit1)

t1_n = 1:12
si1_n=sin(2*pi*w1*t1_n) 

d0_1.new <- forecast(fullfit1, xreg=cbind(si1_n))
autoplot(window(droughtData_1956, start=c(1950,1))) + 
  autolayer(d0_1.new, size=1.5) +
  labs(x="Year", y="Percentage") +
  theme_bw()
```


## Part 2: New Period
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "30"}
#Convert Label

train_2022 %>%
  ggtsdisplay(xlab="Year",
              main="Percentage of US experienced Abnormal Dry to Severe drought 1957-2022")


train_2022 %>% diff() %>%
  ggtsdisplay(xlab="Year",
              main="Differenced Percentage of US experienced Abnormal Dry to Severe drought 1957-2022")
```
### A. Periodogram
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "31"}
f2022 = spec.pgram(train_2022, log="no", spans=c(9,9), detrend = TRUE)
f2022d = spec.pgram(diff(train_2022), log="no", spans=c(9,9), detrend = TRUE)
```
### identify frequency
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "32"}
spx2022 <- f2022d$freq/del
spy2022 <- 2*f2022d$spec
plot(spy2022~spx2022,xlab="frequency",ylab="spectral density",type="l")
#### frequencies: 0.04, 0.19, 0.25, 0.4, 0.5
abline(v=0.05, col="red") 
abline(v=0.155, col="blue")
abline(v=0.275, col="green")
abline(v=0.39, col="pink")
```


### Fit cosine-sine regression model
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "33"}
w1=0.05
w2=0.155
w3=0.275
w4=0.39



t= 1:length(diff(train_2022))
cs1=cos(2*pi*w1*t) 
si1=sin(2*pi*w1*t) 
cs2=cos(2*pi*w2*t) 
si2=sin(2*pi*w2*t)
cs3=cos(2*pi*w3*t) 
si3=sin(2*pi*w3*t) 
cs4=cos(2*pi*w4*t) 
si4=sin(2*pi*w4*t) 
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "34"}
fit2.1=lm(diff(train_2022)~t+cs1+si1+cs2+si2+cs3+si3+cs4+si4) 
summary(fit2.1) 

fit2.2=lm(diff(train_2022)~si1-1) 
summary(fit2.2) 

resi2=fit2.2$residuals
temp2 <- ts(resi2, start=c(1957, 1), frequency=12)
temp2 %>% ggtsdisplay()
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "34 A"}
train_2022 %>% diff() %>% ggtsdisplay()


t2= 1:length(train_2022)
si2=sin(2*pi*w1*t2) 
```

### Fit Arima Model

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "34 B"}
fitp2.1 <- Arima(train_2022, order=c(2, 1, 2), xreg=cbind(si2))
fitp2.2 <- Arima(train_2022, order=c(2, 1, 0), xreg=cbind(si2))
fitp2.3 <- Arima(train_2022, order=c(1, 1, 1), xreg=cbind(si2))
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "34 C"}
tab <- rbind(getAIC(fitp2.1), getAIC(fitp2.2), getAIC(fitp2.3))
tab
colnames(tab) <- c("AIC", "AICc", "BIC")
rownames(tab) <- c("ARIMA(2, 1, 2)",
                   "ARIMA(2, 1, 0)",
                   "ARIMA(1, 1, 1)")
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "34 D"}
tab %>% kbl()%>%
  kable_material(c("striped", "hover"))
```


### Model Accuracy
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "35"}
t2.test = 1:length(val_2022)
si2.t=sin(2*pi*w1*t2.test) 

d0_2.validation <- forecast(fitp2.2, xreg=cbind(si2.t))
acc1 <- accuracy(d0_2.validation, val_2022)
acc1
acc1[2, 2:3]

d0_2.validation2 <- forecast(fitp2.3, xreg=cbind(si2.t))
acc2 <- accuracy(d0_2.validation2, val_2022)
acc2
acc2[2, 2:3]

d0_2.validation3 <- forecast(fitp2.1, xreg=cbind(si2.t))
acc3 <- accuracy(d0_2.validation3, val_2022)
acc3
acc3[2, 2:3]
```

### Visual plot

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "36"}
autoplot(train_2022) +
  autolayer(fitted(fitp2.3))

autoplot(val_2022) + 
  autolayer(d0_2.validation2)

autoplot(window(train_2022, start=c(2015,1))) + 
  autolayer(d0_2.validation2, series="arima with cosine") + 
  autolayer(val_2022, series="True Percentage", size=1.5) +
  labs(x="Year", y="Percentage") +
  theme_bw()
```


### Forecast 3 more data points
```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "37"}
t2_f= 1:length(droughtData_2022)
si2_f=sin(2*pi*w1*t2_f) 


fullfit2 <- Arima(droughtData_2022, order=c(1, 1, 1), xreg=cbind(si2_f))
summary(fullfit2)
checkresiduals(fullfit2)

t2_n = 4:12
si2_n=sin(2*pi*w1*t2_n) 

d0_2.new <- forecast(fullfit2, xreg=cbind(si2_n))
autoplot(window(droughtData_2022, start=c(2000,1))) + 
  autolayer(d0_2.new, size=1.5) +
  labs(x="Year", y="Percentage") +
  theme_bw()
```

# SARIMA Modeling

## SARIMA model for new data

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "38"}
log <- window(train_2022,start=c(1957,1), freq=12)
logD0TS <- log(log)
autoplot(logD0TS)
grid.arrange(ggAcf(logD0TS),ggPacf(logD0TS),nrow=1)
adf.test(logD0TS)

datalog1 <- logD0TS %>% diff(lag=12) 
datalog1 %>% ggtsdisplay()
adf.test(datalog1)

datalog2 <- logD0TS %>% diff(lag=12) %>% diff()
datalog2 %>% ggtsdisplay()
adf.test(datalog2)

log1 <- Arima(logD0TS, order=c(2, 0, 1), seasonal=c(2,0,0))
log2 <- Arima(logD0TS, order=c(1, 0, 1), seasonal=c(1,0,0))
log3 <- Arima(logD0TS, order=c(2, 0, 1), seasonal=c(1,0,0))
log4 <- Arima(logD0TS, order=c(1, 0, 2), seasonal=c(1,0,0))
log5 <- Arima(logD0TS, order=c(2, 0, 2), seasonal=c(1,0,0))
log6 <- Arima(logD0TS, order=c(1, 0, 1), seasonal=c(1,0,1))
log7 <- Arima(logD0TS, order=c(2, 0, 1), seasonal=c(1,0,1))
log8 <- Arima(logD0TS, order=c(1, 0, 2), seasonal=c(1,0,1))
log9 <- Arima(logD0TS, order=c(2, 0, 2), seasonal=c(1,0,1))
log10 <- Arima(logD0TS, order=c(1, 0, 1), seasonal=c(2,0,1))
log11 <- Arima(logD0TS, order=c(2, 0, 1), seasonal=c(2,0,1))
log12 <- Arima(logD0TS, order=c(1, 0, 2), seasonal=c(2,0,1))
log12 <- Arima(logD0TS, order=c(2, 0, 2), seasonal=c(2,0,1))
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "39"}
getAIC <- function(fit) {
  c(fit$aic, fit$aicc, fit$bic)
}

tab <- rbind(getAIC(log1), getAIC(log2), getAIC(log3),getAIC(log4), getAIC(log5), getAIC(log6), getAIC(log7), getAIC(log8), getAIC(log9), getAIC(log10), getAIC(log11), getAIC(log12))
colnames(tab) <- c("AIC", "AICc", "BIC")
rownames(tab) <- c("SARIMA(2,0,1)(2,0,0)","SARIMA(1,0,1)(1,0,0)","SARIMA(2,0,1)(1,0,0)","SARIMA(1,0,2)(1,0,0)",
                   "SARIMA(2,0,2)(1,0,0)","SARIMA(1,0,1)(1,0,1)","SARIMA(2,0,1)(1,0,1)","SARIMA(1,0,2)(1,0,1)",
                   "SARIMA(2,0,2)(1,0,1)","SARIMA(1,0,1)(2,0,1)","SARIMA(1,0,2)(2,0,1)","SARIMA(2,0,2)(2,0,1)")
```

```{r}
tab %>% kableExtra::kbl()%>%
  kable_material(c("striped", "hover"))
```

```{r}
log1
```

$$Y_t(1+\phi_1B)(1+\phi_2B)(1+\Phi_1B^{12})(1+\Phi_2B^{12})= \varepsilon_t(1+\theta_1B)$$
$$Y_t(1+\phi_1B+\phi_2B+\phi_1\phi_2B^2)(1+\Phi_1B^{12}+\Phi_2B^{12}+\Phi_1\Phi_2B^{24})= \varepsilon_t(1+\theta_1B)$$
$$Y_t(1+(\Phi_1+\Phi_2)B^{12}+\Phi_1\Phi_1B^{24}+\phi_1\phi_1B^2+\phi_1\phi_2\Phi_1\Phi_2B^{26}+\phi_1\phi_2(\Phi_1+\Phi_2)B^{14}+ (\phi_1+\phi_2)B+\Phi_1\Phi_2(\phi_1+\phi_2)B^{25}+(\Phi_1+\Phi_2)(\phi_1+\phi_2)B^{13})= \varepsilon_t(1+\theta_1B)$$
$$Y_t = (\phi_1+\phi_2)Y_{t-1} + \phi_1\phi_2Y_{t-2} + (\Phi_1+\Phi_2)Y_{t-12} + (\phi_1+\phi_2)(\Phi_1+\Phi_2)Y_{t-13} + \phi_1\phi_2(\Phi_1+\Phi_2)Y_{t-14} + \Phi_1\Phi_2Y_{t-24} + \Phi_1\Phi_2(\phi_1+\phi_2)Y_{t-25} + \Phi_1\Phi_2\phi_1\phi_2Y_{t-26}+ \varepsilon_t +  \theta_1\varepsilon_{t-1}$$
$$Y_t = 0.962Y_{t-1} -1.325Y_{t-2} -0.02Y_{t-12} -0.019Y_{t-13} + 0.026Y_{t-14} -0.008Y_{t-24} -0.008Y_{t-25} +  0.011Y_{t-26}+ \varepsilon_t -0.717\varepsilon_{t-1}$$

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "40"}
checkresiduals(log1)
Box.test(residuals(log1), lag=4, type="Ljung-Box", fitdf=1)
tsdiag(log1, gof.lag = 15, omit.initial=F)

residualslog <- residuals(log1)
fittedlog <- logD0TS - residualslog
ts.plot(train_2022) + points(exp(fittedlog), type = "l", col = 2, lty = 2)

forecastlog <- forecast(log1, h=15)
forecastlog$lower <- exp(forecastlog$lower)
forecastlog$upper <- exp(forecastlog$upper)
forecastlog$mean <- exp(forecastlog$mean)
forecastlog$x <- exp(forecastlog$x)

drought.predict.test <- ts_reshape(val_2022)
drought.2022 <- drought.predict.test[1:3,c(1,3)]
drought.predict.test <- drought.predict.test[,-c(3)] 
colnames(drought.predict.test)[2] <- "Actual"
drought.predict.test[13:15,] <- drought.2022

forecast.full <- cbind(forecastlog,drought.predict.test[c("Actual")])
forecast.full
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "41"}
acc <- accuracy(forecastlog,val_2022)
acc

autoplot(forecastlog)

autoplot(window(train_2022, start=c(2015,1))) + 
  autolayer(forecastlog, PI=TRUE, series="Forecast") +
  autolayer(val_2022, series="True Drought Percentage", size=1) +
  labs(x="Year", y="Percentage of Country Experiencing Drought") +
  theme_bw()
  
auto.arima(logD0TS)
```

## SARIMA modeling for old data

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "42"}
log2 <- window(train_1956,start=c(1895, 9), freq=12)
logD0TS2 <- log(log2)
autoplot(logD0TS2)
grid.arrange(ggAcf(logD0TS2),ggPacf(logD0TS2),nrow=1)
adf.test(logD0TS2)

datalog12 <- logD0TS2 %>% diff(lag=12) 
datalog12 %>% ggtsdisplay()
adf.test(datalog12)

datalog22 <- logD0TS2 %>% diff(lag=12) %>% diff()
datalog22 %>% ggtsdisplay()
adf.test(datalog22)

log1_2 <- Arima(logD0TS2, order=c(2, 0, 1), seasonal=c(2,0,0))
log2_2 <- Arima(logD0TS2, order=c(1, 0, 1), seasonal=c(1,0,0))
log3_2 <- Arima(logD0TS2, order=c(2, 0, 1), seasonal=c(1,0,0))
log4_2 <- Arima(logD0TS2, order=c(1, 0, 2), seasonal=c(1,0,0))
log5_2 <- Arima(logD0TS2, order=c(2, 0, 2), seasonal=c(1,0,0))
log6_2 <- Arima(logD0TS2, order=c(1, 0, 1), seasonal=c(1,0,1))
log7_2 <- Arima(logD0TS2, order=c(2, 0, 1), seasonal=c(1,0,1))
log8_2 <- Arima(logD0TS2, order=c(1, 0, 2), seasonal=c(1,0,1))
log9_2 <- Arima(logD0TS2, order=c(2, 0, 2), seasonal=c(1,0,1))
log10_2 <- Arima(logD0TS2, order=c(1, 0, 1), seasonal=c(2,0,1))
log11_2 <- Arima(logD0TS2, order=c(2, 0, 1), seasonal=c(2,0,1))
log12_2 <- Arima(logD0TS2, order=c(1, 0, 2), seasonal=c(2,0,1))
log12_2 <- Arima(logD0TS2, order=c(2, 0, 2), seasonal=c(2,0,1))
```

```{r}
getAIC <- function(fit) {
  c(fit$aic, fit$aicc, fit$bic)
}

tab2 <- rbind(getAIC(log1_2), getAIC(log2_2), getAIC(log3_2),getAIC(log4_2), getAIC(log5_2), getAIC(log6_2), getAIC(log7_2), getAIC(log8_2), getAIC(log9_2), getAIC(log10_2), getAIC(log11_2), getAIC(log12_2))
colnames(tab2) <- c("AIC", "AICc", "BIC")
rownames(tab2) <- c("SARIMA(2,0,1)(2,0,0)","SARIMA(1,0,1)(1,0,0)","SARIMA(2,0,1)(1,0,0)","SARIMA(1,0,2)(1,0,0)",
                   "SARIMA(2,0,2)(1,0,0)","SARIMA(1,0,1)(1,0,1)","SARIMA(2,0,1)(1,0,1)","SARIMA(1,0,2)(1,0,1)",
                   "SARIMA(2,0,2)(1,0,1)","SARIMA(1,0,1)(2,0,1)","SARIMA(1,0,2)(2,0,1)","SARIMA(2,0,2)(2,0,1)")
```

```{r}
tab2 %>% kableExtra::kbl() %>%
  kable_material(c("striped", "hover"))
```

```{r}
log1_2
```

$$Y_t = (\phi_1+\phi_2)Y_{t-1} + \phi_1\phi_2Y_{t-2} + (\Phi_1+\Phi_2)Y_{t-12} + (\phi_1+\phi_2)(\Phi_1+\Phi_2)Y_{t-13} + \phi_1\phi_2(\Phi_1+\Phi_2)Y_{t-14} + \Phi_1\Phi_2Y_{t-24} + \Phi_1\Phi_2(\phi_1+\phi_2)Y_{t-25} + \Phi_1\Phi_2\phi_1\phi_2Y_{t-26}+ \varepsilon_t +  \theta_1\varepsilon_{t-1}$$
$$Y_t = 0.96Y_{t-1} -1.273Y_{t-2} -0.009Y_{t-12} -0.009Y_{t-13} + 0.012Y_{t-14} -0.01Y_{t-24} -0.01Y_{t-25} +  0.013Y_{t-26}+ \varepsilon_t -0.7\varepsilon_{t-1}$$

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "44"}
checkresiduals(log1_2)
Box.test(residuals(log1_2), lag=4, type="Ljung-Box", fitdf=1)
tsdiag(log1_2, gof.lag = 15, omit.initial=F)
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "45"}
residualslog2 <- residuals(log1_2)
fittedlog2 <- logD0TS2 - residualslog2
ts.plot(train_1956) + points(exp(fittedlog2), type = "l", col = 2, lty = 2)
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "46"}
forecastlog2 <- forecast(log1_2, h=12)
forecastlog2$lower <- exp(forecastlog2$lower)
forecastlog2$upper <- exp(forecastlog2$upper)
forecastlog2$mean <- exp(forecastlog2$mean)
forecastlog2$x <- exp(forecastlog2$x)
```

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "47"}
drought.predict.test2 <- ts_reshape(val_1956)
colnames(drought.predict.test2)[2] <- "Actual"

forecast.full2 <- cbind(forecastlog2,drought.predict.test2[c("Actual")])
forecast.full2

acc2 <- accuracy(forecastlog2,val_1956)
acc2

autoplot(forecastlog2)

autoplot(window(train_1956, start=c(1950,1))) + 
  autolayer(forecastlog2, PI=TRUE, series="Forecast") +
  autolayer(val_1956, series="True Drought Percentage", size=1) +
  labs(x="Year", y="Percentage of Country Experiencing Drought") +
  theme_bw()
  
auto.arima(logD0TS2)
```

# Final Model Selection 

## Old Data

### Models Selected

List Name of Forecast Selected by Your Process Here

Holt Winters: D0.for.hw.old
Cosine-Sine Model: d0_1.new
SARIMA: forecastlog2

### Model Comparison Table

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "48"}
tab <- rbind(accuracy(D0.for.hw.old, val_1956)[2,2:3],
             accuracy(d0_1.validation, val_1956)[2,2:3],
             accuracy(forecastlog2, val_1956)[2,2:3])
rownames(tab) <- c("Full Holt Winters", "Cosine-sine", "SARIMA")
kable(tab)
```

### Visual Model Comparison

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "49"}
autoplot(window(D0TS.train.old, start=c(1950,1))) + 
  autolayer(D0.for.hw.old, PI=FALSE, series="Full Holt Winters") +
  autolayer(d0_1.validation, PI=FALSE, series="Cosine-sine") +
  autolayer(forecastlog2, PI=FALSE, series="SARIMA") +
  autolayer(D0TS.test.old, series="Real Data") +
  labs(x="Year", y="Percentage of Country Experiencing Drought") +
  theme_bw()
```

### Selection

From a combination of RMSE and MAE, we select the Full Holt Winters model for the overall period from 1895 to 1956.

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "50"}
oldModel <- autoplot(window(D0TS.train.old, start=c(1950,1))) + 
  autolayer(D0.for.hw.old, PI=TRUE, series="Full Holt Winters") +
  autolayer(D0TS.test.old, series="Real Data") +
  labs(x="Year", y="Percentage of Country Experiencing Drought", title = "Old Model") +
  theme_bw()
```

## New Data

### Models Selected

List Name of Model Selected by Your Process Here

Holt Winters: D0.for.season.new
Cosine-Sine Model: d0_2.new
SARIMA: forecastlog

### Model Comparison Table

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "51"}
tab <- rbind(accuracy(D0.for.season.new, val_2022)[2,2:3],
             accuracy(d0_2.validation2, val_2022)[2,2:3],
             accuracy(forecastlog, val_2022)[2,2:3])
rownames(tab) <- c("Holt Season", "Cosine-sine", "SARIMA")
kable(tab)
```

### Visual Model Comparison

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "52"}
autoplot(window(D0TS.train.new, start=c(2017,1))) + 
  autolayer(D0.for.season.new, PI=FALSE, series="Seasonal Only Holt Winters") +
  autolayer(d0_2.validation2, PI=FALSE, series="Cosine-sine") +
  autolayer(forecastlog, PI=FALSE, series="SARIMA(2,0,1)(2,0,0)") +
  autolayer(D0TS.test.new, series="Real Data") +
  labs(x="Year", y="Percentage of Country Experiencing Drought") +
  theme_bw()
```

### Selection

From a combination of RMSE and MAE, we select the SARIMA model for the overall period from 1956 to 2022.

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "53"}
newModel <- autoplot(window(D0TS.train.new, start=c(2017,1))) + 
  autolayer(forecastlog, PI=TRUE, series="SARIMA") +
  autolayer(D0TS.test.new, series="Real Data") +
  labs(x="Year", y="Percentage of Country Experiencing Drought", title = "New Model") +
  theme_bw()
```

# Final Comparison

## Old vs New Visual

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "54"}
grid.arrange(oldModel, newModel, ncol = 2)
```

## Old vs New Mathematical

```{r,cache = TRUE, warning = FALSE, message = FALSE, label = "55"}
val.1956 <- ts_reshape(val_1956)[,-3]
#val.1956[1:12,] <- c(1,2,3,107.6,107.5,113.2) 

retail.full <- cbind(D0.for.hw.old, val.1956[c("1956")])
colnames(retail.full)[6] <- "Actual"
retail.full

val.2022 <- ts_reshape(val_2022)[,-3]
val.2022[13:15,] <- c(c(0,0),c(0,0),c(0,0)) #This is really dumb
#val.1956[1:12,] <- c(1,2,3,107.6,107.5,113.2) 

retail.full2 <- cbind(forecastlog, val.2022)
colnames(retail.full2)[6] <- "Actual"
head(retail.full2,12)
```
